### 2024-02-19 paper_sisc_first_revision/localpsf.tex
done with section 7.1

Here we define the PDE-free spacially varying blurring problem, used
in Sections~\ref{sec:intro} and~\ref{sec:prelims} for visual
illustration of various stages and aspects of the proposed method.  To
study the influence of the presence of negative numbers in the
integral kernel, i.e., the robustness of the proposed method to
violations of the non-negative kernel assumption discussed in
Section~\ref{sec:impulse_response_2}, this example is set up so that
the impulse response, $\phi_x$, corresponding to the integral kernel
is a bumpy blob that is centered near $x$, and is rotated and scaled
in a manner that depends on $x$ (see Figure
\ref{fig:frog_kernel_impulse_responses}). The closed form expression
of the integral kernel is given by
%
%Before delving into technical details, we provide an overview of the
%method. Along the way, we illustrate steps using a spatially variant
%blur example, in which the integral kernel is given as follows:
%\begin{equation}
%\label{eq:frog_kernel}
%\Phi(y,x) = (1 - a f(y,x)) x^1(1-x^1)x^2(1-x^2) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).
%\end{equation}
%Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation
%matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2
%\operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and
%$f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right)
%\sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a,
%L, c_1, c_2$. In this section we use $a=1.0$, $L^2=1.0$, $c_1=0.01$,
%and $c_2=0.0025$.
%
\begin{align}
\label{eq:example_frog_kernel}
\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}(h(y,x)^T
C^{-1} h(y,x) \right),
\end{align}
where $f(y,x) = \cos\left(h^1(y,x) / \sqrt{c_1 /
  2}\right)\sin\left(h^2(y,x) / \sqrt{c_2 / 2}\right)$, with
$h^i(y,x)$ the $i^{th}$ component of $R(\theta(x)) (y-x)$, with
$R(\theta(x))$ a two-dimensional rotation matrix by angle $\theta(x) =
(x^1 + x^2) \pi / 2$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and $C = L^2
\operatorname{diag}(c_1, c_2)$. The constant $L$ controls how wide the
blob is (hence the support of the impulse response), $a$ controls the
magnitude of the bumps, and $(c_1, c_2)$ control the width of the
Gaussian. We note that when $a=0$, $\phi_x$ is a Gaussian, as $a$
increases, $\phi_x$ becomes more bumpy. For instance, when $a > 1$,
the integral kernel contains negative values, which violates the
Assumption 3 (see Section~\ref{sec:conditions_2}).

@@

Define HODLR. Define randomized HODLR algo. Ref to Tucker's paper and "Fast randomized eigensolver with structured LDL..."

Define randomized SVD, refer to Halko Martinsson Tropp.

Discuss why PSF does better with smaller $L$.

\begin{table}
	\begin{center}
	{\small
	\begingroup
	\setlength{\tabcolsep}{8pt}
	\renewcommand{\arraystretch}{1.0}
	\begin{tabular}{c| c | c c c c c}
				& Error tol.    &	\#applies PSF   & \#applies HODLR 	& \#applies RSVD 	\\%& rank \\ 
	\hline 
				& 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
	$L=1$ 		& 10\% 			& 15  				& 772				& 520				\\%& 178  \\ 
				& 5\% 			& 23   				& 924				& 674				\\%& 244 \\ 
	\hline
				& 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
	$L=1/2$ 	& 10\%    		& 10  				& 1144				& 1916				\\%& 665 \\ 
				& 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 
	\hline
				& 20\%    	 	& 7					& 932				& 2624				\\%& 845  \\ 
	$L=1/3$ 	& 10\%    		& 8  				& 1264 				& 3734				\\%& 1321  \\ 
				& 5\% 			& 8   				& 1520 				& 4660				\\%& 1759  \\ 
	\end{tabular}
	\endgroup
	}
	\end{center}
	\caption{(Spatially variant blur) Comparison of cost to approximate the blur kernel from Equation \eqref{eq:frog_kernel} using the PSF method, the randomized HODLR (hierarchical off diagonal low rank) method, and GLR (global low rank) approximation using randomized SVD. The quantity $L$ scales the width of the impulse responses, hence it influences the rank of the operator. Large $L$ means low rank, and small $L$ means high rank. The second column (``Error tol'') is the relative error in the approximation of the kernel measured in the Frobenius norm, $||\Phi - \widetilde{\Phi}||_\text{Fro} / ||\Phi||_\text{Fro}$. The remaining three columns show the number of operator applies required to achieve the given error tolerances, using the PSF, HODLR, and GLR methods.}
	\label{tab:frog_psf_vs_hodlr_vs_glr}
\end{table}

>>>>

Here we define a PDE-free spatially varying blurring problem, used in
Sections~\ref{sec:intro} and~\ref{sec:prelims} for visual illustration
of various stages and aspects of the proposed method.  To study the
robustness of the proposed method to violations of the non-negative
kernel assumption discussed in Section~\ref{sec:impulse_response_2},
this example is set up so that the impulse response, $\phi_x$,
corresponding to the integral kernel is a bumpy blob that is centered
near $x$, and is rotated and scaled in a manner that depends on $x$
(see Figure \ref{fig:frog_kernel_impulse_responses}). The closed form
expression of the integral kernel is given by
%
%Before delving into technical details, we provide an overview of the
%method. Along the way, we illustrate steps using a spatially variant
%blur example, in which the integral kernel is given as follows:
%\begin{equation}
%\label{eq:frog_kernel}
%\Phi(y,x) = (1 - a f(y,x)) x^1(1-x^1)x^2(1-x^2) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).
%\end{equation}
%Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation
%matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2
%\operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and
%$f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right)
%\sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a,
%L, c_1, c_2$. In this section we use $a=1.0$, $L^2=1.0$, $c_1=0.01$,
%and $c_2=0.0025$.
%
\begin{align}
\label{eq:frog_kernel}
\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}(h(y,x)^T
C^{-1} h(y,x) \right),
\end{align}
where $f(y,x) = \cos\left(h^1(y,x) / \sqrt{c_1 /
  2}\right)\sin\left(h^2(y,x) / \sqrt{c_2 / 2}\right)$, with
$h^i(y,x)$ the $i^{th}$ component of $R(\theta(x)) (y-x)$, with
$R(\theta(x))$ a two-dimensional rotation matrix by angle $\theta(x) =
(x^1 + x^2) \pi / 2$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and $C = L^2
\operatorname{diag}(c_1, c_2)$. The constant $L$ controls the width of
the blob, and $c_1/c_2$ controls its aspect ratio. The constant $a$
represents deviation from a Gaussian. When $a=0$, $\phi_x$ is a
Gaussian, and as $a$ increases, $\phi_x$ becomes non-Gaussian. When $a
> 1$, the integral kernel contains negative values, which allows us to
study the robustness of the PSF method to violations of Assumption 3
(Section~\ref{sec:conditions_2}).

@@

In Table~\ref{tab:frog_psf_vs_hodlr_vs_glr} we compare the
computational cost to approximate the blur kernel from
Equation~\eqref{eq:frog_kernel} using the proposed PSF method, the
randomized HODLR (hierarchical off diagonal low rank)
method~\cite{Martinsson16,HartlandEtAl23}, and GLR (global low rank)
approximation using randomized SVD~\cite{HalkoMartinssonTropp11}. For
these results we vary the quantity $L$ to scale the width of the
impulse responses and hence the rank of the operator. For each case,
we calculate the relative error in the approximation of the kernel
measured in the Frobenius norm, $||\boldsymbol{\Phi} -
\widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
||\boldsymbol{\Phi}||_\text{Fro}$ and show the number of operator
applications required to achieve 20\%, 10\%, and 5\% relative error by
each method. The results reveal that superior performance of the PSF
method as compared to the HODLR and GLR methods for all cases. We note
that as we increase the rank and decrease the error tolerance, the
performance of the the HODLR and GLR methods deteriorates.

%Define HODLR. Define randomized HODLR algo. Ref to Tucker's paper and "Fast randomized eigensolver with structured LDL..."

%Define randomized SVD, refer to Halko Martinsson Tropp.

%Discuss why PSF does better with smaller $L$.

\begin{table}
	\begin{center}
	{\small
	\begingroup
	\setlength{\tabcolsep}{8pt}
	\renewcommand{\arraystretch}{1.0}
	\begin{tabular}{c| c | c c c c c}
				& Error tol.    &	\#applies PSF   & \#applies HODLR 	& \#applies RSVD 	\\%& rank \\ 
	\hline 
				& 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
	$L=1$ 		& 10\% 			& 15  				& 772				& 520				\\%& 178  \\ 
				& 5\% 			& 23   				& 924				& 674				\\%& 244 \\ 
	\hline
				& 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
	$L=1/2$ 	& 10\%    		& 10  				& 1144				& 1916				\\%& 665 \\ 
				& 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 
	\hline
				& 20\%    	 	& 7					& 932				& 2624				\\%& 845  \\ 
	$L=1/3$ 	& 10\%    		& 8  				& 1264 				& 3734				\\%& 1321  \\ 
				& 5\% 			& 8   				& 1520 				& 4660				\\%& 1759  \\ 
	\end{tabular}
	\endgroup
	}
	\end{center}
	\caption{(Spatially variant blur) Comparison of cost to
          approximate the blur kernel from Equation
          \eqref{eq:frog_kernel} using the PSF method, the randomized
          HODLR (hierarchical off diagonal low rank) method, and GLR
          (global low rank) approximation using randomized SVD. The
          quantity $L$ scales the width of the impulse responses,
          hence it influences the rank of the operator. Large $L$
          means low rank, and small $L$ means high rank. The second
          column (``Error tol'') is the relative error in the
          approximation of the kernel measured in the Frobenius norm,
          $||\boldsymbol{\Phi} -
          \widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
          ||\boldsymbol{\Phi}||_\text{Fro}$. The remaining three
          columns show the number of operator applies required to
          achieve the given error tolerances, using the PSF, HODLR,
          and GLR methods.}
	\label{tab:frog_psf_vs_hodlr_vs_glr}
\end{table}
